%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
% This program is free software; you can redistribute it and/or modify %
% it under the terms of the GNU General Public License as published by %
% the Free Software Foundation; either version 2 of the License, or    %
% (at your option) any later version.                                  %
%                                                                      %
% This program is distributed in the hope that it will be useful,      %
% but WITHOUT ANY WARRANTY; without even the implied warranty of       %
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the        %
% GNU General Public License for more details.                         %
%                                                                      %
% You should have received a copy of the GNU General Public License    %
% along with this program; if not, write to the Free Software          %
% Foundation, Inc., 51 Franklin St, Fifth Floor, Boston,               %
% MA  02110-1301  USA                                                  %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%	$Id$
%

\begin{thebibliography}{99}
	\bibitem{reg-exp}Jeffrey E. F. {\sc Friedl}
		\textsl{Mastering Regular Expressions,
				Powerful Techniques for Perl and Other Tools} --
		1st Edition January 1997 -- O'Reilly \& Associates Inc.

	\bibitem{sed-awk}Dale {\sc Dougherty} \& Arnold {\sc Robbinsx},
		\textsl{Sed \& Awk Programming} --
		2nd Edition March 1997 -- O'Reilly \& Associates Inc.

	\bibitem{learning-emacs}Debra {\sc Cameron}, Bill {\sc Rosenblatt} \& Eric {\sc Raymond},
		\textsl{Learning GNU Emacs} --
		2nd Edition September 1996 -- O'Reilly \& Associates Inc.

	\bibitem{ksh}Bill {\sc Rosenblatt},
		\textsl{Learning the Korn Shell} --
		June 1993 -- O'Reilly \& Associates Inc.

	\bibitem{learning-perl}Randal L. {\sc Schwartz}  and Tom {\sc Christiansen},
		\textsl{Learning Perl} --
		Second edition, July 1997 -- O'Reilly \& Associates Inc.

	\bibitem{programming-perl}Larry {\sc Wall}, Tom {\sc Christiansen} and Randal L. {\sc Schwartz},
		\textsl{Programming Perl} --
		Second edition, September 1996 -- O'Reilly \& Associates Inc.

	\bibitem{advpgm-perl}Sriram {\sc Srinivasan},
		\textsl{Advanced Perl Programming} --
		March 1992 -- O'Reilly \& Associates Inc.

	\bibitem{dns-bind}Paul {\sc Albitz} and Cricket {\sc Liu},
		\textsl{DNS and Bind} --
		Mars 1993 -- O'Reilly \& Associates Inc.

	\bibitem{int-ip}Sylvain {\sc Baudry}.
		\textsl{Les r{\'e}seaux 802.3 et IP, Gestion avec SNMP} --
		Juin 1993 -- RNUR.

	\bibitem{admcent}Sylvain {\sc Baudry}.
		\textsl{Administration centralis{\'e}e de plusieurs {\Unix}s} --
		Septembre 1994 -- INED.

	\bibitem{nfs-nis-automnt}Sylvain {\sc Baudry}.
		\textsl{Description des protocoles NFS et NIS -- Applications et
			administration sur diff{\'e}rents {\Unix}} --
		Avril 1995 -- INED.

	\bibitem{bsd-sock}Sylvain {\sc Baudry}.
		\textsl{Programmation r{\'e}seau avec les BSD-Sockets} --
		Juin 1996 -- ESME.

	\bibitem{guide-latex}Helmut {\sc Kopka} and Patrick W. {\sc Daly}.
		\textsl{A Guide to \LaTeX2e{}, Document Preparation for Beginners and Advanced Users} --
		Second edition, 1995 -- Addison-Wesley

	\bibitem{langage-rationnel}\textsl{Langage rationnel} --
		\href{http://fr.wikipedia.org/wiki/Langage_rationnel}{Wikipedia}

	\bibitem{expression-rationnelle}\textsl{Expressions rationnelles} -- 
		\href{http://fr.wikipedia.org/wiki/Expression_rationnelle}{Wikipedia}

\end{thebibliography}
